---
title: "Reproducible Research: Peer Assessment 1"
author: "Olks"
date: "15 listopada 2015"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
Loading dataset and calculating sums:

```{r, echo=TRUE}
activitySet <- read.csv("activity.csv")
library(dplyr)
sumByDay <- activitySet %>% 
                group_by (date) %>% 
                summarize (sumByDay=sum(steps,na.rm=T))
```


Histogram of sums of steps per day:

```{r, echo=TRUE}
library(ggplot2)
qplot(sumByDay$sumByDay)
```

## What is mean total number of steps taken per day?
Calculating mean and median:

```{r, echo=TRUE}
meanSteps <- mean(sumByDay$sumByDay,na.rm=T)
medianSteps <- median(sumByDay$sumByDay,na.rm=T)
```

Mean of the total number of steps taken per day is eqial `r meanSteps`.
Median of the total number of steps taken per day `r medianSteps`.

## What is the average daily activity pattern?
Calculating average steps in each interval across all the days in the dataset:

```{r, echo=TRUE}
meanByInterval <- activitySet %>% 
                        group_by (interval) %>% 
                        summarize (meanByInterval=mean(steps,na.rm=T))

ggplot(meanByInterval, aes(interval, meanByInterval)) + geom_line() +
        xlab("5 min intervals during the day") + ylab("Mean number of Steps")

intervalNr <- meanByInterval$interval[which.max(meanByInterval$meanByInterval)]
```

The maximum number of steps, on average across all the days in the dataset,
contains the interval `r intervalNr`.

## Imputing missing values
Filling NAs and making histogram of sums of steps per day:

```{r, echo=TRUE}

nrofNAs <- sum(!is.na(activitySet$steps))

fullSet <- activitySet
for (i in 1:length(fullSet$steps)){
        if (is.na(fullSet$steps[i])){
             interNr <- fullSet$interval[i]
             fullSet$steps[i] <- 
                     subset(meanByInterval,interval==interNr)$meanByInterval
        }
}


sumByDayFull <- fullSet %>% 
        group_by (date) %>% 
        summarize (sumByDayFull=sum(steps,na.rm=T))

library(ggplot2)
qplot(sumByDayFull$sumByDayFull)
```


## Are there differences in activity patterns between weekdays and weekends?
Calculating differeces between dataset with NAs and without NAs:

```{r, echo=TRUE}
meanStepsFull <- mean(sumByDayFull$sumByDayFull,na.rm=T)
medianStepsFull <- median(sumByDayFull$sumByDayFull,na.rm=T)
```

Mean for oryginal dataset: `r meanSteps`.
Mean for complete dataset: `r meanStepsFull`.

Median for oryginal dataset: `r medianSteps`.
Median for complete dataset: `r medianStepsFull`.



Creating two plots for weekdays and weekends:

```{r, echo=TRUE}
fullSet$date<-as.Date(fullSet$date)
fullSet$weekday<-as.integer(format(fullSet$date, "%w"))

fullSet$week<-0
fullSet[fullSet$weekday>5,]$week<-1
fullSet$week<-factor(fullSet$week,labels=c("weekday","weekend"))

meanByInt <- fullSet %>% 
        group_by (interval,week) %>% 
        summarize (meanInt=mean(steps,na.rm=T))



qplot(interval, meanInt, facets=week~., data=meanByInt,geom=c("line"),
      ylab="number of steps")
```


